### 2015-08-20 manuscript_supportinginfo.Rmd

---
title: Supporting Information
author: ''
date: ''
output:
  word_document:
    reference_docx: manuscriptstyle.docx
  pdf_document: default
  html_document:
    number_sections: yes
csl: global-change-biology.csl
bibliography: references.bib
---

```{r echo=FALSE, warning=FALSE, message=FALSE}
source("manuscript_calcs.R")
opts_chunk$set(dpi=300)
```

## Supporting Figures

```{r fig.width=4, fig.height=6, echo=FALSE, message=FALSE, eval=TRUE}
figureSI1(litring, subset(facegap_cloudy_byring, Date < maxdate))
```

**Figure S1**. Increase in gapfraction ($\tau_d$) and total litter production during the 4-month drought period in 2013, calculated for each ring separately. Blue colours are ambient (nrs 2,3,6), red colours elevated C~a~ (nrs 1,4,5). (a) The change in  $\tau_d$ and its 95% confidence interval was calculated from the slope of linear regression of $\tau_d$ with time (number of days since 8 July 2013). (b) Total litter production during the drought expressed in terms of *L* for each ring. The confidence interval was calculated based on the 8 litter traps in each ring. Note the overlapping confidence intervals for all rings.

\
```{r fig.width=8, fig.height=5, echo=FALSE, message=FALSE, eval=TRUE}
figureSI3(subset(facegap_cloudy_byring, Date < maxdate), 
          subset(facegap_all_byring, Date < maxdate))
```

**Figure S2.** Comparison of canopy transmittance ($\tau$) during highly diffuse conditions (F~diff~ > 0.98), or using all PAR data without filtering. Shown are daily average $\tau$ for all six FACE rings (symbols), with a smoother (generalized additive model with 95% confidence intervals). Note the much lower $\tau$ during southern hemisphere winter when using all data compared to diffuse only. This difference arises due to low solar angles, which lead to higher attenuation of the direct beam as a result of a longer path length through the canopy.

\
```{r fig.width=8, fig.height=4, echo=FALSE, message=FALSE, eval=TRUE}
figureSI2(flatcan_byring)
```

**Figure S3**. Correlation between *L* estimated from diffuse canopy transmittance ($L_\tau$) and from canopy photos (L~p~). Solid line is a 1:1 line, dashed line a regression line across all data (with the shading indicating the 95% confidence interval), and the colored lines regressions per plot across dates. (a) Estimates of L~p~ with the extinction coefficient (*k*) set to 0.5, as recommended by MacFarlane et al. [-@Macfarlane2007] ($L_\tau$ = `r tidy(lmcanphot)$estimate[1]` + `r tidy(lmcanphot)$estimate[2]` L~p~, R^2^ = `r glance(lmcanphot)$r.squared`, p `r pval(glance(lmcanphot)$p.value)`), (b) Estimates of L~p~ with *k* optimized to minimize the difference between L~p~ and $L_\tau$ ($L_\tau$ = `r tidy(lmcanphot2)$estimate[1]` + `r tidy(lmcanphot2)$estimate[2]`).

\
```{r fig.width=4, fig.height=4, echo=FALSE, message=FALSE}
figureSI4(ba)
```

**Figure SI4**. Average $LAI_\tau$ over the entire study period plotted against the basal area (sum of stem cross-sectional areas) of the FACE rings ($R^2$ = `r glance(balm)$r.squared`, *L* = `r tidy(balm)$estimate[1]` + `r tidy(balm)$estimate[2]` BA). This relationship was used to calculate the *L* anomaly in Fig. 3, as the difference between *L* predicted from this regression and the actual $LAI_\tau$ at any given time. Red symbols are elevated C~a~, blue symbols ambient C~a~.

\
```{r fig.width=8, fig.height=4, echo=FALSE, message=FALSE}
res <- figureSI5(df, ba)
```

**Figure SI5**. Detectable differences in *L* between C~a~ treatments after accounting for pre-treatment variation in basal area. We fit *L* with a fitted GAM using C~a~, time and pre-treatment basal area (BA) as the covariates. In the GAM, time was entered as a smooth term (estimated by a cubic spline), and BA as a linear term. Shown are the predictions (smooth lines) from the fitted GAM when BA was set to the mean value across all plots. Shaded areas indicate 95% confidence intervals (CI). Note they are much narrower than the CI across plots when BA was allowed to vary (Fig. 2 in main text). The CI here thus accounts for the pre-treatment differences between rings, and visualizes how using BA as a covariate in the mixed-effects model increases the ability to resolve small difference in *L* that may develop over time. The minimum detectable difference in *L* between C~a~ treatments can be estimated from the average half-width of the CI of fitted *L* at a common BA. A significant effect would could be observed (provided the error variance remains the same) when the two CIs (for aC~a~ and eC~a~) would just not overlap (so that p < 0.05), or twice the half-width CI. This difference was estimated as `r round(res$halfci,2)` m^2^ m^-2^, which when expressed relative to the mean *L* of `r round(mean(ba$LAI.mean),2)` is a difference of only `r round(100*res$minobsdiff,1)`%.

## Notes S1. Estimation of leaf area index from digital photography

We took advantage of the detailed timeseries of *L* estimates from canopy transmittance to test a simple photography-based method proposed by Macfarlane et al. [-@Macfarlane2007], which is extensively applied to *Eucalyptus*-dominated forests. From October 2012 to October 2013, we took ca. monthly canopy photographs in each FACE ring. Photographs were taken in the morning on cloudy days. We took one photograph at each point of a 4 x 4 m grid established within each ring, unless obstructed by trees, yielding 20-23 photos per ring. Sampling points too close to either the vertical vent pipes (edges of the ring) or the central scaffolding tower were excluded. The camera was aligned with the north-south axis and set at 1.25 m height.  The camera lens was pointed directly upwards and leveled using a three axis bubble level. We used a Nikon D90 camera with a Nikon AF 50 mm lens (Nikon Corporation, Tokyo, Japan). To measure cover, the camera was set to automatic exposure, Aperture-Priority and automatic autofocus mode and F10 lens. Following recommendations by Jonckheere et al. [-@Jonckheere2005], we used the automated Ridler algorithm to threshold the canopy photos. We then calculated canopy transmittance ($\tau_p$) from each canopy photo as the proportion of pixels with intensity in the blue band below that threshold. *L* from canopy photos (*L*~p~) was calculated with a simple Beer-Lambert law (Eq. 4).

\
(4) 
$$L_p = -log(\tau_p)/k$$

where $\tau_p$ canopy transmission estimated from canopy photos. We assumed that the extinction coefficient *k* = 0.5, following Macfarlane et al. [-@Macfarlane2007] recommendation for eucalypt forests. Without further calibration, the two estimates were reasonably well correlated, though on average *L*$_\tau$ was `r f(100*(with(flatcan_byring, mean(LAI.PAR.mean / LAI))-1))`% higher than $L_p$ (R^2^ = `r f(glance(lmphoto)$r.squared)`, RMSE = `r f(glance(lmphoto)$sigma)`) (Fig. S3). We also estimated the extinction coefficient that minimized the difference between the two estimates of *L* (Fig. S2), giving *k* = `r f(kopt)`. These comparisons are encouraging because the photos measure *L* near the zenith (+/- 15^o^), while $L_\tau$ is integrated over the sky hemisphere (Eq. 2), weighted towards the zenith.

## References