### 2015-08-21 manuscript.Rmd
Edited according to co-authors comments on revision; suppinfo has own template

---
title: Canopy leaf area of a mature evergreen *Eucalyptus* woodland does not respond to elevated atmospheric [CO~2~] but tracks water availability
author: ''
date: ''
output:
  word_document:
    reference_docx: manuscriptstyle.docx
  pdf_document: default
  html_document:
    number_sections: yes
csl: global-change-biology.csl
bibliography: references.bib
---

\
*Running head* : Leaf area index in elevated [CO~2~]

\
Remko A. Duursma^1^, Teresa E. Gimeno^1,2^, Matthias M. Boer^1^, Kristine Y. Crous^1^, Mark G. Tjoelker^1^, David S. Ellsworth^1^.

\
^1^ Hawkesbury Institute for the Environment, University of Western Sydney, Locked Bag 1797, Penrith, NSW, Australia

\
^2^ INRA, UMR 1391 ISPA, 33140 Villenave d'Ornon, France

\
*Corresponding author*: 
Remko Duursma
\
E: remkoduursma@gmail.com
\
T: +61(0)422096908

\
*Keywords*: 
leaf area index, FACE, phenology, drought, litter production

\
*Type of Paper*:
Primary Research Article

```{r echo=FALSE, message=FALSE}
source("manuscript_calcs.R")

# Make figures
# source("make_figures.R")

includeFigs <- FALSE
```

# Abstract {.unnumbered}

Canopy leaf area, quantified by the leaf area index (*L*), is a crucial driver of forest productivity, water use, and energy balance. Because *L* responds to environmental drivers, it can represent an important feedback to climate change, but its responses to rising atmospheric [CO~2~] and water availability of forests have been poorly quantified. We studied canopy leaf area dynamics for `r nummon` months in a native evergreen *Eucalyptus* woodland exposed to Free Air CO~2~ Enrichment (the EucFACE experiment), in a subtropical climate where water limitation is common. We hypothesized that, because of expected stimulation of productivity and  water-use efficiency, *L* should increase with elevated [CO~2~]. We estimated *L* from diffuse canopy transmittance, and measured monthly leaf litter production. Contrary to expectation, *L* did not respond to elevated [CO~2~]. We found that *L* varied between `r f(rangeLAI[1])` and `r f(rangeLAI[2])` across the study period. The dynamics of *L* showed a quick increase after heavy rainfall, and a steady decrease during periods of low rainfall. Leaf litter production was correlated to changes in *L*, both during periods of decreasing *L* (when no leaf growth occurred), and during periods of increasing *L* (active shedding of old foliage when new leaf growth occurred). Leaf lifespan, estimated from mean *L* and total annual litter production, was up to two months longer under elevated [CO~2~] (`r f(LLtest$estimate2)` vs. `r f(LLtest$estimate1)` yrs; p `r pval(LLtest$p.value)`). Our main finding that *L* was not responsive to elevated CO~2~ is consistent with other forest FACE studies, but contrasts with the positive response of *L* commonly predicted by many ecosystem models.

# Introduction

Leaf area index (*L*) is an important ecosystem attribute underlying productivity of vegetation [@Reich2012], and a key variable in land surface models [@Kala2013; @Walker2014]. *L* is a dynamic canopy property, as it is generally responsive to variation in water and nutrient availability, as demonstrated in experiments [@Herbert1995; @Hubbard2004; @White2010] and observational studies across rainfall gradients [@Ellis2008; @Iio2013; @Jin2013].  Ecosystem models generally predict that *L* increases with elevated atmospheric [CO~2~] (further denoted as eC~a~), largely in response to increased net carbon uptake [@DeKauwe2014; @Hatton1992; @Kergoat2002; @Zhu2012]. This potential change in *L* is of central importance for the net effect of eC~a~ on water, CO~2~ and energy fluxes between vegetation and atmosphere as it may modify or even reverse the direct physiological responses to eC~a~ [@Tor-ngern2015; @Murray2012]. Despite its importance, little empirical evidence is available on the response of *L* to eC~a~ in mature forests [@Norby2011].

\
Increased carbon availability through enhanced photosynthesis under eC~a~ does not necessarily lead to increases in *L*. Direct effects of eC~a~ on vegetation processes include an enhancement of photosynthesis and productivity, a decrease in stomatal conductance and leaf-level water use and an increase in water-use efficiency [@Ainsworth2005; @Ellsworth2012]. Although numerous studies have examined the response of physiology and plant growth to eC~a~, very few of these inform the response of steady-state *L* to eC~a~, because it requires a sufficiently mature forest to be exposed to Free Air CO~2~ Enrichment (FACE) [@Norby2011]. Out of the five forest FACE studies that have been carried out worldwide, only broad-leaved deciduous aspen stands (in monoculture or mixed with birch; Rhinelander-FACE) and needle-leaved evergreen pine stands (DukeFACE) showed increased *L* in response to eC~a~ [@McCarthy2007; @Norby2011], although positive responses to eC~a~ in terms of productivity were observed at all five sites [@Norby2005]. Interestingly, Norby and Zak [-@Norby2011] found a negative correlation between pre-treatment *L* and the responsiveness to eC~a~, leading to the prediction that stands with low *L* will show a larger response to eC~a~.

\
In ecosystems where water availability limits productivity, we expect a larger response of *L* to eC~a~ [@Donohue2009]. This is because eC~a~ leads to reduced stomatal conductance, and hence water use per unit leaf area [@Warren2010] leading to increased soil water availability. These 'water savings' can be expected to lead to an increase in *L* in water-limited ecosystems, consistent with a tight coupling of *L* with precipitation across gradients [@Ma2013; @Jin2013; @Ellis2008]. In line with this prediction, Donohue et al. [-@Donohue2013] attributed the observed global increase in *L* in arid and semi-arid areas over the last 30 years to rising atmospheric [CO~2~]. Based on the same water-savings principle, we predicted that *L* increases under eC~a~ in water-limited forests.

\
The responses of *L* to fluctuations in water availability are qualitatively well-known [@Pook1985; @Specht1989; @Whitehead2004; @Pook1997; @Limousin2009], but poorly quantified. The strong relationship between *L* and rainfall observed across Australia [@Ellis2008; @Specht1989; @Broich2014], and other continents [@Donohue2013; @Grier1977; @Jin2013] represents the adaptation of vegetation structure and density to long-term climate conditions. However, local *L* is also highly responsive to water availability at shorter time scales [@Pook1985; @Pook1997], driven by the dynamics of leaf turnover [@Misson2011] and leaf flushing and expansion [@Muller2011]. Understanding dynamics of *L* in response to water availability and its potential interactions with eC~a~ is crucial if we are to improve phenology models in water-limited ecosystems [@Misson2011].

\
We estimated *L* from canopy transmittance in a *Eucalyptus*-dominated mature woodland exposed to Free-Air CO~2~ Enrichment (FACE), the EucFACE experiment. The strong seasonal and interannual variability in precipitation at the site allowed us to quantify the dynamics of *L* and its components (litter fall and production) in response to water availability. We hypothesize that eC~a~ would result in an increase in *L*, particularly during periods of water limitation. In addition, we expect that the low *L* that is typical of native *Eucalyptus*-dominated woodlands [@Whitehead2004] allow a positive response of *L* to eC~a~ either via an increase in the rate of leaf production or leaf retention (leaf lifespan). Our second hypothesis is that the temporal dynamics of *L* and its components (litter fall and leaf production), are related to to fluctuating water availability.

# Materials and Methods

## Site description and experimental design
The EucFACE experiment is located in a remnant patch  of native Cumberland Plain woodland [@Tozer2003] on an ancient alluvial floodplain in western Sydney (Australia, 33^o^ 37'S, 150^o^ 44'E, 30 m a.s.l.). The site is characterized by a humid temperate-subtropical transitional climate with a mean annual temperature of 17 ^o^C and mean annual precipitation (P) of 800 mm (1881 - 2014, Bureau of Meteorology, station 067105 in Richmond, NSW Australia; http://www.bom.gov.au). Estimated potential evapotranspiration (PET) is 1350 mm [1950 - 2000 climatology, http://www.cgiar-csi.org, @Zomer2008], giving a ratio of P/PET of ca. 0.59. In addition, over the last 30 years, monthly PET has exceeded monthly P 78% of the time, indicating frequent water limitation.

\
At the site, the soil at the surface is a loamy sand with > 75% sand content, and from ca. 50 to 300 cm depth, soils are a sandy clay loam with > 30% silt and clay. The soil is nutrient poor, and a strongly P-limited [@Crous2015]. Groundwater depth is 12.6 - 12.9 m below the soil surface. The vegetation consists of an open woodland (600-1000 trees ha^-1^, basal area before treatment, BA = `r f(mean(ba$BA))`, SE = `r f(sd(ba$BA)/sqrt(6))` m^2^ha^-1^, n = 6 plots), with a single dominant species (*Eucalyptus tereticornis* Sm.) and some scattered individuals of *E. amplifolia* Naudin.  The understory stratum is dominated by grasses, with lower densities of forbs and occasional shrubs.

\
Within the EucFACE site, there are six 25-m diameter circular plots (designated hereafter as 'rings'). Each ring comprises 32 vent pipes in a cylindrical frame extending above the tree canopy to 28 m high. Each vertical pipe is subtended by a control valve and is connected to a circular plenum [@Hendrey1999] through which a computer-controlled system modulates release of diluted CO~2~ into the vegetation [@Lewin2009]. The vegetation within three randomly selected rings has been exposed to a CO~2~ mole fraction higher than ambient since September 2012, whereas the other three rings were identical to the CO~2~ treatment rings in infrastructure and instrumentation but received aC~a~, and are used as control plots. The target daytime C~a~ for the rings exposed to eC~a~ was 150 $\mu$mol CO~2~ mol^-1^ above aC~a~ in full operation mode. However, C~a~ was increased gradually for six months at a rate of ca. 30 $\mu$mol CO~2~ mol^-1^ month^-1^ until full operation mode and the target eC~a~ was reached in February 2013 (Fig. 1a, [@gimeno2015]).

\
Air temperature was measured on top of the central tower of each ring (23.5 m height) with a HUMICAP sensor (HMP 155 Vaisala, Vantaa, Finland). Total precipitation was monitored with three automated tipping buckets (Tipping Bucket Rain gauge TB4, Hydrological Services Pty Ltd, Liverpool, NSW, Australia) located at the top of the central tower at 23.5 m height, in three of the rings. Total precipitation and mean temperature were logged every 15 minutes onto CR3000 data loggers (Campbell Scientific, Townsville, Australia). Soil volumetric water content was measured at two different locations in each ring with two sets of theta probes (ThetaProbe ML2x, Delta-T Devices Ltd, Cambridge, UK) installed at 5, 30, 75 cm depth and at the depth at which a hard pan with higher clay content was present (30-50cm depth depending on the ring).

## Terminology
In the following, we use the term leaf area index (*L*) although the canopy and FACE rings contain non-leaf shading elements, i.e. woody stems and infrastructure. This avoids the need to switch between different terminologies when comparing temporal changes in *L* to litter production, and absolute values between different methods. The temporal changes in *L* are due to leaf area dynamics only, because infrastructure area is constant, and variation in woody stem area can be assumed negligible since no tree mortality has occurred over the two-year study period.

## Leaf litter production
In each ring, eight litter traps were installed at randomly selected locations. The litter traps were 0.1979 m^2^ each with a depth of 0.4 m, and sat at 0.5m above the forest floor. Litter traps were custom-made from plastic baskets coated with a fine nylon mesh with a 200 $/mu$m mesh size. To minimize contamination with litter from trees outside the rings, the baskets were installed at least 2 m from the plot edge. Litter was collected monthly since October 2012, and leaves were separated from other components such as twigs, bark, seeds, insect frass, etc. Litter production was converted to units of *L* by multiplying it with an estimate of specific leaf area (SLA) determined from 329 live, non-senescent leaves collected from the tree canopies from May 2012 to May 2014 (`r get_sla()` cm^2^ g^-1^, SE=2.14). We did not find a significant effect of C~a~ treatment on SLA (Ellsworth, unpublished data) on any of the collection dates.

## *L* from diffuse transmittance
In each of the six FACE rings, three photosynthetically active radiation (PAR) sensors (LI-190, Li-cor, Lincoln, NE, U.S.A.) were installed on metal posts at 1 m height. In each ring, another PAR sensor was installed above the canopy, on a mast extending above a scaffolding tower in the centre of each ring at 23.5 m height. Data were logged on data loggers (CR3000, Campbell Scientific) at one minutely intervals beginning on 15 October 2012, and automatically uploaded every day to a data management system.  All PAR data were averaged at 30-min time steps. At a nearby flux tower site (within 2 km), we measured the diffuse fraction (F~diff~) of PAR at 30-min time step using a BF5 Sunshine sensor (Delta-T Instruments, Cambridge, UK) installed on a tower extending ca. 5 m above the forest canopy. For determination of *L*, we used only highly diffuse conditions, where F~diff~ > 0.98. We further trimmed the data by excluding low solar elevation (< 10^o^), because measured canopy transmittance was more variable at these low solar angles. Finally canopy transmittance was calculated for every 30-min time step as,

\
(1) 
$$\tau_d = \bar{Q}_{below} / Q_{above}$$

where $\tau_d$ is diffuse canopy transmittance, Q~above~ is above canopy PAR ($\mu$mol m^-2^s^-1^), and Q~below~ is the average PAR across the three sub-canopy PAR sensors in each FACE ring. We then calculated daily averages of $\tau_d$, excluding days where fewer than 4 time steps (i.e. 2 hours) were available, and days where the standard deviation among the three PAR sensors in a ring was > `r .maxSD` (in units of $\tau_d$).

\
We then estimated leaf area index (*L*) from $\tau_d$ following Campbell and Norman [-@Campbell2000]. We assumed that leaves had a spherical leaf angle distribution, and that leaves were randomly distributed in space. We recognize that these assumptions are very likely violated, but we account for this bias by calibrating the *L* estimate to leaf litter production as further detailed below. Diffuse transmittance of a canopy is a function of the *L* by integrating the Beer-Lambert law across the entire sky hemisphere, and accounting for differences in extinction coefficient with different viewing angles (because of leaf angle and longer path length through the canopy at low solar angles).

\
(2)
$$\tau_d = 2\int_0^{\pi/2} e^{-LG(\theta)/cos(\theta)}sin(\theta)cos(\theta)d\theta$$

where $\theta$ is the solar zenith angle, G($\theta$) is the foliage projection function in the direction of the beam [@Campbell2000]. Because we assumed a spherical leaf angle distribution, G($\theta$) = 0.5 [@Stenberg2006]. We solved numerically for *L* in Eq. 2, using a smooth integration over $\theta$, yielding *L*.

\
To overcome violations in the assumptions underlying Eq. 2, we calibrated *L* against cumulative litter production during a 4-month drought period in 2013 (indicated in Fig. 1). During this period (14 July 2013 to 12 Nov 2013), no leaf growth occurred and *L* decreased steadily (and thus $\tau_d$ increased steadily, see Fig. 1). During this period total precipitation was 42mm, the second driest July-October period recorded at a nearby (<2km) weather station since 1881 (compared to a long-term average of 188 mm). A total of n = 11 days with sufficiently diffuse conditions to estimate $\tau_d$  were available during this period. Cumulative *L* leaf litter production (expressed in units of *L*) should match the observed decrease in *L* estimated from canopy transmittance. We calibrated Eq. 3,

\
(3) 
$$c(L_{\tau,t} - L_{\tau,t=0}) = \int_0^t P_Ldt$$

where *c* is a calibration parameter, $L_{\tau,t=0}$ and $L_{\tau,t=1}$ are the *L* at the start and the end of the drought period, respectively. We used linear regression without an intercept to estimate *c* in Eq. 4 as `r f(calib)` (SE = `r f(tidy(caliblm)$std.error)`). It should be stressed that even after calibration with litter production, *L* still includes stem and infrastructure area (see Terminology). However, this approximately constant area cancels out in the left-hand side of Eq. 3. We used a single calibration constant for all rings, because we did not find significant differences between the rings in terms of either litter production or change in $\tau_d$ during the 2013 drought period (see Fig. S1).

\
We used the opportunity provided by the high quality timeseries of *L* to test a simple photography-based method suggested by Macfarlane et al. [-@Macfarlane2007]. The two methods compared quite well (R^2^ = `r f(glance(lmcanphot)$r.squared)`) (Notes S1, Fig. S3), suggesting both diffuse transmittance and the photography-based method can be reliably used in *Eucalyptus*-dominated woodlands to estimate *L*.

## Data analysis
We used generalized additive models (GAM) for all analyses of *L* and $\tau_d$ timeseries [@Wood2006], using the *mgcv* package in R version `r r_version` [@RCT2015]. A GAM is a semi-parametric fitting method that allows the estimation of smoothed trends in the data, including confidence intervals. In all fitted GAMs, we used a cubic regression spline. For the smoothed term in the model (days since start of experiment), we used up to `r .gambase` degrees of freedom, which resulted in biologically realistic smoothed *L* dynamics. Within the GAM, we used a penalized regression smoother [@Wood2006] to allow the final degree of smoothness to be estimated from the data. Random effects were included to account for the repeated measures structure of the data, except when we analysed average *L* across all rings.

\
To test for effects of C~a~ treatment on *L* and litter production, we used linear-mixed effects models with FACE ring as the random effect, and C~a~ treatment and date as factor variables. The random effect accounts for the repeated measures nature of the data. For *L*, we further used pre-treatment basal area (BA) of each FACE ring as a covariate (see Fig. S4). Because the EucFACE woodland is well established, the differences in BA are not due to C~a~ treatment effects, but instead represent natural variation within the woodland.

\
To compare the dynamics of *L* with litter production, we calculated the change in *L* for every time period when litter was collected. We did this by fitting a GAM to *L* for each ring separately (similar to that shown for $\tau_d$ in Fig. 1), and calculating the change in *L* for each litter fall collection period ($\Delta$*L*). Next, we estimated gross leaf production (G~L~) as the sum of the net change in *L* ($\Delta$*L*) and P~L~, because mass balance requires that $\Delta$*L* = G~L~ - P~L~. We estimated mean leaf lifespan (LL, yr) as the ratio of mean *L* (m^2^ m^-2^) for the entire litter fall collection period to total P~L~ (m^2^ m^-2^ yr^-1^).

\
We quantified the relationship between timing of leaf flushing and environmental drivers. During the `r nummon` month study period, we observed four leaf flushing events in the field, with one before estimates of *L* were available. These events occurred in spring and summer, with no clear seasonal cycle (October 2012, January 2013, December 2013 and January 2015). These leaf growth events were compared to the dynamics of *L* as estimated from a GAM fitted to the average daily *L* across the six FACE rings. From the fitted model, we estimated the numerical derivative of *L* (dL/dt), as well as its confidence interval following Simpson [-@Simpson2014, see also @Curtis2014].

# Results
Canopy transmittance ($\tau_d$) varied between `r f(min(df$Gapfraction.mean))` and `r f(max(df$Gapfraction.mean))`, with much temporal variation, and substantial absolute differences between the individual FACE rings (Fig. 1).  However, temporal trends were very similar between the rings. Since the understorey comprises mostly grasses and forbs, and the PAR sensors were installed at 1 m height, this variation is due to the overstorey trees only. An outbreak of psyllids (*Cardiaspina* sp.), a tiny sap-feeding insect, resulted in rapid loss of canopy leaf area in one of the rings (ring 6), starting in September 2014 [for description of a similar outbreak in *Eucalyptus moluccana*, see @Hall2015; pers. comm. Andrew Gherlenda, Markus Riegler]. This rapid decrease in *L* was evident in a quick increase in $\tau_d$ (Fig. 1). However, it recovered rapidly after rainfall in December 2014, resulting in a lower $\tau_d$ than before the insect attack.

\
We found no differences in *L* between the C~a~ treatments, as shown by consistent overlap of the 95% confidence intervals of the generalized additive model (thus, p > 0.05 at any point in time, Fig. 2). A linear mixed-effects model with C~a~, time and pre-treatment basal area (Table 1) as fixed effects, and FACE ring as a random effect also showed no significant effect of C~a~ on *L* (p > 0.5) (Table 1).

\
Litter production (P~L~) correlated significantly with $\Delta$*L* when the latter was separated into periods where $\Delta$*L* > 0 (net leaf growth) and $\Delta$*L* <  0 (net leaf loss) (Fig. 3). During periods of net leaf loss, we found a significant correlation (with a slope of `r f(tidy(lmlitdown)$estimate[2])`, SE = `r f(tidy(lmlitdown)$std.error[2])`, n = `r glance(lmlitdown)$df.residual+2`, p < 0.001), close to the 1:1 line, as expected since no leaf growth was observed in the field during any period where *L* was decreasing (see also Fig. 6). During periods of net leaf growth, we found a significant correlation (with a slope of `r f(tidy(lmlitup)$estimate[2])`, SE = `r f(tidy(lmlitup)$std.error[2])`, n = `r glance(lmlitup)$df.residual+2`, p < 0.001) between $\Delta$*L* and litter production (Fig. 4), indicating that most shedding of old foliage occurred during periods of active leaf growth.

\
The temporal dynamics of litter and leaf production are shown in Fig. 4, and demonstrate the irregular temporal pattern in both leaf production and P~L~. We found no significant difference in either litter or leaf production between C~a~ treatments at any time point (p > 0.2 for both, assessed with linear-mixed effects models).

\
Next, we related total P~L~ to average *L* across the six FACE rings (Fig. 5a). If leaf lifespan (LL) is constant, P~L~ would be linearly related to *L*. Instead, we found that in eC~a~, P~L~ was lower at a given *L* as compared to aC~a~ (Fig. 5a). We found a significant (p `r pval(LLtest$p.value)`) increase in LL with eC~a~ (Fig. 5b), with LL = `r f(LLtest$estimate1)`yr in aC~a~ and `r f(LLtest$estimate2)`yr in eC~a~.

\
The dynamics of *L* from the sensor record matched leaf growth events observed in the field quite well (Fig. 6a). All four leaf flushing events corresponded to significant positive dL/dt. One brief significant positive dL/dt (June 2013) did not correspond to an observed leaf flushing event, but it was small in magnitude. It is also obvious that every leaf flush event occurred shortly after a significant rainfall event (Fig. 6b,c), when the soil water content in the top 80 cm increased to ca. > 0.2 m^3^ m^-3^. Two periods of sustained decrease in *L* (red lines in Fig. 6b, during winter of 2013 and 2014) corresponded to periods of low soil water availability.

# Discussion

## Response of *L* to elevated [CO~2~]
We report on `r nummon` months of continuous leaf area index (*L*) measurements in a native evergreen *Eucalyptus* woodland exposed to +150ppm elevated CO~2~ (eC~a~). We predicted that *L* would increase in response to eC~a~ as a result of enhanced CO~2~ assimilation for the entire canopy and increased soil water availability. We also expected that this eucalypt woodland, with its inherently low *L* [@Whitehead2004], would be even more responsive to eC~a~ than higher-*L* and more closed-canopy stands [@Norby2011]. Counter to our expectations but similar to Norby et al. [-@Norby2003], we found no significant effect of eC~a~ on *L*. Because estimated leaf lifespan (LL) averaged `r round(mean(ba$LL),1)` yrs (Fig. 5), the evergreen canopy has regrown at least twice during the study period, giving ample opportunity for eC~a~ effects to occur. It could be argued that the change in *L* in response to eC~a~ was too small to be detected. Natural variation in *L* among plots in native woodland is indeed a challenge to manipulative studies with low replication, such as FACE experiments. To increase sensitivity and statistical robustness, we used pre-treatment stand basal area as a covariate. Using this method, we estimated that we can detect a difference as little as 6.1% in *L* (see Fig. S5). Based on this small detectable difference, we conclude that the lack of a significant C~a~ effect is not due to a lack of power, but a true lack of response to eC~a~.

\
This lack of a response to eC~a~ counters our expectations and those in previous publications [@Norby2003; @McCarthy2007] for a number of reasons. First, in a review of all forest FACE sites, Norby and Zak [-@Norby2011] showed that low *L* stands were more responsive to eC~a~ than stands with high *L*. They argued that high *L* canopies would not benefit from further increases in *L*, because it would lead to only very small increases in total absorbed radiation. The mean *L* at the EucFACE was ca. `r f(mean(ba$LAI.mean))` m^2^m^-2^, which is low compared to a global compendium of forest *L* [@Iio2013], and the lowest *L* of any forest FACE experiment [@Norby2011], but in line with a range of values for native *Eucalyptus* forests and woodlands reported by Whitehead and Beadle [-@Whitehead2004] and Iio et al. [-@Iio2013]. Based on this *L* estimate and the relationship across FACE sites reported by Norby and Zak [-@Norby2011], we expected about a 30% increase in *L* under eC~a~. In contrast to other forest FACE sites, the EucFACE is strongly limited by nutrient availability, in particular phosphorus [@Crous2015; @Nielsen2015]. Hence, it may be too early to draw general conclusions across FACE sites in terms of *L* response.

\
Stimulation of net leaf photosynthesis by eC~a~ (a common response, also observed at the EucFACE, Gimeno et al. [-@gimeno2015]) may not lead to increased net leaf growth, due to other constraints such as water or nutrient limitations. Net leaf growth is the net sum of allocation of photosynthates to leaf growth, and turnover of the leaf pool. Either process may itself be affected by eC~a~ [@DeKauwe2014]. We found a significant effect of eC~a~ on LL, with leaves living approximately `r round(abs(LLtest$estimate*12),1)` months longer in eC~a~ (Fig. 5) (an increase of `r round(100*with(LLtest,estimate2/estimate1 - 1),0)`%). Because we found no difference in *L* with C~a~ treatment, this suggests a compensating difference in the fractional allocation of net primary production to leaf growth. To further understand the response of biomass allocation to eC~a~ at the EucFACE, all components of the carbon balance must be carefully quantified, a task underway but not yet complete.

\
Despite a lack of response of canopy *L* to eC~a~, we did find a significantly longer LL in eC~a~. We could not determine whether there was an interaction between LL and water availability, since we estimated leaf turnover from mean *L* and total accumulated leaf litter (Fig. 5). Although LL was increased in eC~a~, this was not reflected in a difference in leaf litter production at any time point (Fig. 4), suggesting the change in LL reflects a small but significant shift in leaf retention, not related to decreased litter production during drought. Previous results from other forest FACE studies are not conclusive in this regard. Norby et al. [-@Norby2003] reported inconsistent effects of eC~a~ on leaf area duration in a deciduous *Liquidambar* plantation (ORNL-FACE). At the same site, Warren et al. [-@Warren2011] reported increased senescence in the eC~a~ treatment in response to a combined drought and heat wave. McCarthy et al. [-@McCarthy2007] found no effects of eC~a~ on phenology in either *Pinus* sp. or *Liquidambar* sp. canopies at the DukeFACE, despite enhancement of *L* by eC~a~. Taylor et al. [-@Taylor2008] reported delayed autumn senescence in *Populus* sp. in two FACE experiments (PopFACE and AspenFACE). Although this indicates the canopy stayed green longer in eC~a~, Gielen et al. [-@Gielen2001] reported faster turnover rate of leaves in eC~a~ at the PopFACE, leading to similar *L* in aC~a~ despite higher productivity. Clearly, no consistent picture emerges from these comparisons.

## Litter production and dynamics of *L*
We found a correlation between *L* dynamics and litter production, when the data were analyzed separately for periods of decreasing and increasing *L* (Fig. 3). As expected, during periods of decreasing *L*, the change in *L* was correlated with litter production, and we were able to use a four-month period in 2013 without significant rainfall to calibrate estimates of *L* against litter production. During periods of increasing *L*, litter fall increased sharply (Fig. 3), indicating replacement of old leaves by young leaves. Indeed we observed in the field that very few old leaves were present in the canopy after the significant flushing event of December 2013. This is in line with Pook [-@Pook1984], who showed in *Corymbia maculata* (a genus closely related to *Eucalyptus*) accelerated litter production during periods of net *L* increase, but with a lag between litter fall and production of about 2 months. Similarly, Stone and Bacon [-@Stone1995] showed that drought-stressed trees had higher litter production at the branch scale, but irrigated trees had larger litter production at the canopy scale. Since leaf flushing has been shown to be closely related to soil water availability in eucalypts [Fig. 6; @Pook1984a], net leaf turnover is driven by water availability both through increased litter production during extended droughts, and complete canopy replacement following leaf flushing (Fig. 4). Given the high inter-annual variation in rainfall at the EucFACE site, leaf turnover is highly variable in time.

## Continuous measurement of *L*
Studying canopy leaf area index in an evergreen native woodland poses unique challenges. Unlike in deciduous canopies, litter production cannot be used to estimate *L*, because LL is unknown (instead, we used *L* and litter production to estimate LL). We could also not estimate leaf area from allometric relationships with stem diameter or height, since these equations are not available for mature *E. tereticornis*, and would not be valuable anyway since *L* is temporally variable (Fig. 6). We found that indirect estimates of *L* based on canopy transmittance were very reliable because relatively small temporal changes in *L* could be detected (see Fig. S5), indicating leaf flushing events after rainfall, increased litter production during drought, and an insect attack (Figs. 1, 4, 6). We do argue that the raw data must be filtered by diffuse conditions, as the spatial and within-day variance are negligible compared to those during periods of direct sunshine [@Chason1991; @Dufrene1995; @Nilson1999]. As expected, when we used unfiltered canopy transmittance data, the seasonal patterns in canopy transmittance were much more distinct (Fig. S3), because longer path lengths in the winter result in lower overall transmittance. Using diffuse PAR alone avoids having to correct for these path length effects.

## Consequences for phenology models
Current-generation phenology models are heavily biased towards temperate conditions and deciduous rather than evergreen species [@Primack2011], where leaf flushing occurs in spring in response to increase in temperature [@Richardson2012; @Savoy2015]. In contrast, phenology of evergreen *Eucalyptus* is more opportunistic, and shows rapid response to water availability (Fig. 6). Specht and Brouwer [-@Specht1975] demonstrated for four *Eucalyptus* species that leaf growth  peaked both in spring and autumn, with low winter temperatures and high summer temperatures (and low water availability) suppressing leaf growth. Similarly, Pook  [-@Pook1984a] demonstrated occurrence of leaf production during most months in *Corymbia maculata*, except during winter and during dry periods. Thus there seems to be a minimum temperature threshold for *Eucalyptus* leaf growth. Similarly, we also have not observed leaf flushing in winter during the study period, but at our site low temperatures were confounded with water availability as both winters during the study period were relatively dry (Fig. 6).

\
Although the response of *L* to water availability is qualitatively well known, current-generation phenology models fall short in incorporating this response. In a comparison of six leading ecosystem and global vegetation models,  Powell et al. [-@Powell2013] showed that the steady decrease in *L* in an Amazon drought experiment could be matched by only one model. Simulations at the EucFACE site using the set of ecosystem models in De Kauwe et al. [-@DeKauwe2014] also showed little response to inter-annual variation in rainfall (Medlyn et al. *in review*). In most models  that provide prognostic estimates of *L* [i.e., excluding models that take *L* as an input parameter; e.g. @Kala2013], temporal changes in *L* are modelled by first calculating total net photosynthetic productivity (carbon availability, gross photosynthesis minus respiration), and then allocating some fraction to the leaf pool. Combined with - usually assumed constant - leaf turnover and specific leaf area, this yields predictions of *L* in response to environmental drivers such as eC~a~. We argue that this algorithm is not well suited to explain the dynamic nature of *L* in response to water availability observed in evergreen broad-leaved forests. In order to construct a phenology model where phenology is responsive to water availability, it should include responses of both leaf turnover and leaf growth to fluctuating water availability, as we have demonstrated that both processes are highly dynamic. Water limitation is known to directly increase leaf turnover [@Pook1997], possibly to maintain sufficient water availability to the remaining leaves [@Vilagrosa2003; @Misson2011], and decrease rates of leaf expansion [e.g. @Muller2011]. Some existing ecosystem models allow for increased leaf turnover during periods of low water availability (e.g. G'DAY, see Corbeels et al. 2005), but this process has been poorly quantified. Our study demonstrates that both responses of leaf turnover and leaf growth to fluctuating water availability are important to improve predictions of seasonal dynamics of *L* in ecosystem models.

# Acknowledgements {.unnumbered}
We thank Steven Wohl, Vinod Kumar, Craig Barton and Craig McNamara for maintaining the FACE site and its wide array of sensors, and Andrew Gherlenda, Cassie Mosdal, Tamara Leetham and Catherine Beattie are thanked for significant efforts in separating and weighing the litter samples. We also thank Gerry Devine for support in data management. EucFACE is an initiative supported by the Australian Government through the Department of Industry and Science in partnership with the University of Western Sydney. Facilities and instrumentation at EucFACE were assembled as an initiative of the Australian Government as part of the Nation-building Economic Stimulus Package. Additional support from the Australian Research Council is acknowledged. The entire analysis presented in this paper will be made available as a *git* repository (http://www.github.com/remkoduursma/XXX), published on figshare (doi:XXX). The repository includes the raw data needed to run the analyses, the source code for the analysis and manuscript (which was written in markdown, using the 'rmarkdown' R package for conversion).

# Tables {.unnumbered}

**Table 1**. Results of a mixed-effects model of *L* with C~a~, Date and pre-treatment basal area (BA) as predictors. The model accounts for repeated measures by using FACE ring as the random effect. Sample size of *L* per ring was `r table(df$Ring)[1]`.

```{r results='asis', echo=FALSE}
lmer_LAI[,"p-value"] <- formatPval(lmer_LAI[,"p-value"])
lmer_LAI[,"F"] <- round(lmer_LAI[,"F"],2)
kable(lmer_LAI)
```

# Figures {.unnumbered}

```{r, echo=FALSE, message=FALSE}
options(scipen = 1, digits = 2)
```

```{r fig.width=8, fig.height=6, echo=FALSE, message=FALSE, eval=includeFigs}
figure1(subset(facegap_cloudy_byring,Date < maxdate), ramp)
```

**Figure 1**. (a) Target [CO~2~] (C~a~) relative to the aC~a~ treatment at the EucFACE. The eC~a~ treatment proceeded in steps of 30 ppm, reaching ambient +150 ppm in February 2013. (b) Canopy transmittance ($\tau_d$) estimated from continuous time series of photosynthetically active radition (PAR) below and above the canopy, for highly diffuse conditions (F~diff~ > 0.98) only. The solid lines are smoothed time series (with 95% confidence intervals) for each FACE ring (using a generalized additive model). Lines in blue colours denote aC~a~ treatment, red colours are eC~a~. Note the steep increase in $\tau_d$ for Ring 6 around November 2014, due to an insect attack that resulted in near defoliation of the canopy. The arrow labelled 'calibration' indicates a drought period when $\tau_d$ steadily increased, used to calibrate *L* estimates to litter production (see Methods).

```{r fig.width=8, fig.height=4, echo=FALSE, message=FALSE, eval=includeFigs}
figure2(subset(facegap_cloudy_byring, Date < maxdate))
```

**Figure 2**. Time series of *L* estimated from canopy transmittance. Each symbol is a value for a FACE ring, but the smooth curve (fitted with a generalized additive model) takes into account the random plot effect of the three replicate rings. Note that the 95% confidence always overlap, denoting no significant eC~a~ treatment effect at $\alpha$ = 0.05. Variation between rings was substantial, and largely due to pre-treatment differences in stand basal area (see Fig. S4).

```{r fig.width=6, fig.height=4, echo=FALSE, message=FALSE, echo=FALSE, message=FALSE, eval=includeFigs}
figure3(dLAIlitter)
```

**Figure 3**. Comparison of absolute changes in *L* ($\Delta$*L*) to litter production (P~L~), calculated for each litter collection period. Green symbols (circles) indicate periods during which $\Delta L$ was positive (net leaf production), orange symbols (triangles) when $\Delta$*L* was negative (net leaf loss). Solid black lines are 1:1 lines, coloured lines regression lines with 95% confidence intervals.

```{r fig.width=8, fig.height=4, echo=FALSE, message=FALSE, eval=includeFigs}
figure4(dLAIlitter)
```

**Figure 4**. Timeseries of litter production (P~L~) and estimated leaf production rate. Leaf production rate was estimated as the sum of the net change in *L* and litter production (converted to a per area basis using specific leaf area estimates, see Methods). Error bars are omitted for clarity. We did not find any differences in P~L~ or leaf production between C~a~ treatments (see Results).

```{r fig.width=8, fig.height=4, echo=FALSE, message=FALSE, eval=includeFigs}
figure5(ba)
```

**Figure 5**.  (a) Total P~L~ (converted from data in Fig. 4 to an annual estimate) for each of the FACE rings plotted against mean *L* during the entire litter collection period. The ratio of mean *L* to annual litter production gives the mean leaf life span. Grey lines indicate leaf lifespan isolines (in years). (b) Calculated leaf lifespan for the six FACE rings shows a significant difference between C~a~ treatments (t-test, p `r pval(LLtest$p.value)`).

```{r fig.width=10, fig.height=8, echo=FALSE, message=FALSE, eval=includeFigs}
figure6(subset(facegap_cloudy_byring, Date < maxdate),
               subset(facesoilwater, Date < maxdate),
               subset(faceraindaily, Date < maxdate),
               subset(airt, Date < maxdate))
```

**Figure 6**. Dynamics of *L* in relation to environmental drivers. (a) Smoothed time series of *L* averaged across all six FACE rings. (b) The daily change in *L* (dL/dt), including the 95% confidence interval. Green indicates significant positive dL/dt, red indicates significant negative dL/dt (at $\alpha$ = 0.05). Four of the five significant positive increases corresponded to approximate dates of commencement of leaf flushing (horizontal lines) observed in the field (except the third brief significant increase). (c) Time course of soil water content measured over 0-80 cm soil depth interval, averaged across all sensors and all rings (blue line), and daily rainfall (black lines).  (d) Average daily air temperature with a fitted smoother.

# Supporting information captions {.unnumbered}

**Figure S1**. Increase in gapfraction ($\tau_d$) and total litter production (P~L~) during the 4-month drought period in 2013, calculated for each ring separately.

**Figure S2**. Correlation between *L* estimated from diffuse canopy transmittance and from canopy photos.

**Figure S3**. Comparison of canopy transmittance ($\tau$) during highly diffuse conditions (F~diff~ > 0.98), or using all PAR data without filtering.

**Figure S4**. Average *L* over the entire study period is correlated with pre-treatment basal area (sum of stem cross-sectional areas) of the FACE rings.

**Figure S5**. Detectable differences in *L* between C~a~ treatments after accounting for pre-treatment variation in basal area.

# References {.unnumbered}